### 2013-11-28 manuscripts/reply/reply-figures.tex
Working on reply ms.
More work on the section about empirical Bayes model choice theory.

\mFigure{../images/coin_flip.pdf}{
    A plot of three beta probability density functions that represent a prior
    (black; $beta(0.5, 0.5)$), true posterior (blue; $beta(4.5, 1.5)$), and
    empirical Bayes density (red; $beta(8.5, 2.5)$) for a dataset of five
    Bernoulli trials, four of which are successes.
}{figCoinFlip}

>>>>

\mFigure{../images/coin_flip.pdf}{
    A plot of three beta probability density functions that represent a prior
    (black; $beta(10, 10)$), true posterior (blue; $beta(13, 17)$), and
    empirical Bayes density (red; $beta(16, 24)$) for a dataset of 10 Bernoulli
    trials, three of which are successes.
}{figCoinFlip}

### 2013-11-28 manuscripts/reply/reply-main.tex
Working on reply ms.
More work on the section about empirical Bayes model choice theory.

Whereas empirical Bayes approaches can provide powerful methods for parameter
estimation, a theoretical justification for empirical Bayes approaches to model
choice is questionable.
In Bayesian model choice, the primary goal is not to estimate parameters, but
to estimate the relative probabilities of candidate models.
In a simple example where two candidate models, $M_1$ and $M_2$, are being
compared, the goal is to estimate the posterior probabilities of these two
models.
Again, we can use Bayes' rule to calculate this as
\begin{equation}
    p(M_1 \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given
    M_1)p(M_1)}{p(\alignment{i}{} \given M_1)p(M_1) + p(\alignment{i}{} \given
    M_2)p(M_2)}.
    \label{eq:bayesmodelchoice}
\end{equation}
By comparing Equations \ref{eq:bayesrule} and \ref{eq:bayesmodelchoice}, we
see fundamental differences between Bayesian parameter estimation and
model choice.
In Equation \ref{eq:bayesrule}, we see that the posterior density of any
state $\myTheta{}$ of the model, is the prior density updated by the
probability of the data given the state $\myTheta{}$ (the likelihood of
$\myTheta{}$).
The integral over the entire parameter space of the model, which is defined by
the priors, only appears as a normalizing constant in the denominator.
Thus, as long as the joint prior distribution contains the values of
$\myTheta{}$ that maximize the probability of the data (i.e., the maximum
likelihood) and the data are strongly informative relative to the prior, the
modes of the estimated posterior distributions of parameters will be relatively
robust to prior choice, even if the shape of the posterior is incorrect due to
using the data to inform the priors.
This is why empirical Bayes can work well for estimating the values of model
parameters.

@@

However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
Bayesian model choice it is now the \emph{marginal} likelihood of a model that
updates the prior to yield the model's posterior probability.
Thus the integral over the entire parameter space of the likelihood weighted by
the prior probability density is no longer a normalizing constant, but rather
is what informs the posterior probability of that model from the data.
As a result, Bayesian model choice tends to be much more sensitive to priors
than Bayesian parameter estimation.
In other words, the Bayesian model choice is much more sensitive to the marginal
likelihood of models than parameter estimation.
Whereas parameter estimation will tend to succeed if the topography of the
likelihood surface is captured regardless of scale, model choice must compare
the average height of this surface between models.

Another important difference of Bayesian model choice illustrated by
Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the
posterior probability of a model, is not a function of \myTheta{}, because
it is integrated out of the marginal likelihoods of the candidate models.
Thus, unlike parameter estimates, the estimated posterior probability of a
model is a single value (rather than a distribution) lacking a measure of
posterior uncertainty.
Also, unlike model parameters, the posterior probabilities of candidate
models have no clear true values.
In most applications, we know all candidate models are incorrect, but our goal
is to summarize the relative probabilities that they could produce the observed
data.
Thus, model posterior probabilities are inherently measures of our belief in the
models after our prior beliefs are updated by the data being analyzed.

This complicates the meaning of model posterior probabilities when Bayes' rule
is violated by empirically informing priors.
For empirical Bayesian parameter estimation, the objective is that by giving the
data more weight relative to the prior, the posterior distribution will be
peaked near the true value(s) of the model's parameter(s).
There is no such justification for empirical Bayesian model choice, because
there are no true values for the model probabilities being estimated.
By using the data twice, we fail to account for prior uncertainty and mislead
our posterior beliefs in the models being compared; we will be over confident
in some models and under confident in others.

These distinctions between Bayesian parameter estimation and model choice can
be illustrated with a simple example.
Let us say we are interested in the fairness of a particular coin, and we
denote the probability of landing heads as \myTheta{}.
More specifically, we are interested in the probability of two models, $M_1$
and $M_2$.
In both models the outcomes of flipping the coin are assumed to be binomially
distributed, but under $M_1$ the coin is weighted toward landing heads (i.e.,
$\myTheta{} > 0.5)$), where as under $M_2$, the coin is weighted toward landing
tails (i.e., $\myTheta{} < 0.5$).
We already have data from flipping a different coin 20 times that landed both
heads and tails 10 times, and so we decide to use these data in specifying a
beta prior on fairness of the new coin of $beta(a=10, b=10)$
(Figure~\ref{figCoinFlip}).
We collect data by flipping the coin $N=10$ times, $y=3$ of which land heads.
Given the beta distribution is a conjugate prior for a binomial likelihood, the
posterior distribution has the nice analytical form $\theta \given y,N \sim
beta(a + y, b + N - y)$, which for the new dataset is simply $beta(13, 17)$
(Figure~\ref{figCoinFlip}).
The posterior mode estimate of the probability of heads is 0.429, and following
Equation~\ref{eq:marginallikelhiood} the marginal likelihoods of our models of
interest are
\begin{equation}
    p(y=3, N=10 \given M_1) = \int_{0.5}^{1} p(y=3, N=10 \given
    \myTheta{}, M_1)p(\myTheta{} \given M_1) d\myTheta{} \approx 0.029,
\end{equation}
and
\begin{equation}
    p(y=3, N=10 \given M_2) = \int_{0}^{0.5} p(y=3, N=10 \given
    \myTheta{}, M_2)p(\myTheta{} \given M_2) d\myTheta{} \approx 0.097.
\end{equation}
Given the models have equal probability under our prior, and following
Equation~\ref{eq:bayesmodelchoice}, we can calculate the posterior
probability of Model 1 as
\begin{equation}
    p(M_1 \given y=3, N=10) = \frac{p(y=3, N=10 \given M_1)}{p(y=3, N=10 \given
    M_1) + p(y=3, N=10 \given M_2)} \approx 0.23.
\end{equation}
This is the true posterior probability of Model 1 given our prior information
and data.

To give the data more weight relative to the prior, we could use it twice, and
calculate an empirical Bayes estimate using a prior of $beta(13,17)$.  This
results in a distribution of $beta(16, 24)$ (Figure~\ref{figCoinFlip}), with a
mode of 0.395, and $p(M_1 \given y=3, N=10) = 0.10$.
Whereas the estimate of the parameter $p(heads)$ is a distribution, the mode of
which changed only 8\% under the empirically informed prior, the posterior
probability of Model 1 is a single value, which changed by 56\% under the
empirical prior.
This demonstrates how Bayesian model choice is more sensitive to priors than
parameter estimation.
Also, we can see how using the empirically informed prior ignores our prior
uncertainty, leading to an underestimate of our posterior uncertainty
(Figure~\ref{figCoinFlip}), and ultimately misleads us toward overconfidence in
Model 2.
Our mode estimate of \myTheta{} under the empirical prior might be closer
to the true value than our posterior mode.
Let us assume the truth of the matter is that the coin under study has a broad
flat edge, and as a result lands on its edge at a certain frequency.
Hence, both of our models are incorrect (i.e., the outcomes of flipping the
coin are not binomially distributed).
From this example, we can see how Bayesian model choice is more sensitive to
priors, and how using the data twice can mislead the posterior probability of a
model in light of new data.

In addition to these theoretical concerns there are practical problems with
using narrow, empirical, uniform priors for a method that has already been
shown to be biased toward models with less parameter space.

@@

% \cwlnote{}{It seems like this section can be boiled down to ``empirical Bayes
%     approaches for model choice when using uniform priors can exclude the
%     $truth$ from the prior, which is a violation of Bayes theorem" or something
%     like that. Picking apart what they calculated and we calculated seems like
%     too much detail. I would try to simplify how this is being presented. This
%     can lead into how the use of more appropriate prior distributions should be
%     used.}
The results of Hickerson et al.'s \citeyear{Hickerson2013} reanalysis of the
Philippine dataset strongly favored models with the narrowest, empirically
informed prior on divergence time, and thus their model-averaged posterior
estimates are dominated by models $M_1$ and $M_2$ (see Table 1 of
\citet{Hickerson2013}).
This is concerning, because the narrowest \divt{} prior used by
\citet{Hickerson2013} ($\divt{} \sim U(0,0.1)$) likely excludes the true
divergence times for at least some of the Philippine taxa, a major problem when
using uniform priors.
\citet{Hickerson2013} set this prior to match the 95\% highest posterior
density (HPD) interval for the mean divergence time estimated under one of the
priors used by \citet{Oaks2012} (see Tables 2 and 3 of \citet{Oaks2012}).
Given this interval estimate is for the \emph{mean} divergence time across all
22 taxa, it may be inappropriate to set this as the limit on the prior, because
many of the taxon pairs are expected to have diverged at times older than the
upper limit.
Furthermore, this prior is \emph{excluded} from the 95\% HPD interval estimates
of the mean divergence time under the other two priors explored by
\citet{Oaks2012} (under these priors the 95\% HPD is approximately 0.3--0.6;
see Table~6 of \citet{Oaks2012}).
Thus, the results of \citet{Hickerson2013} indicate that integration over
parameter space with low likelihood is biasing the method toward models with
smaller parameter space, and as a consequence is biasing the method toward
models that exclude the truth (i.e., toward estimating model-averaged
posteriors dominated by models that exclude true values of parameters of the
model).

>>>>

Whereas empirical Bayes approaches can provide powerful methods for parameter
estimation, a theoretical justification for empirical Bayes approaches to model
choice is questionable.
In Bayesian model choice, the primary goal is not to estimate parameters, but
to estimate the relative probabilities of candidate models.
In a simple example where two candidate models, $M_1$ and $M_2$, are being
compared, the goal is to estimate the posterior probabilities of these two
models.
Again, we can use Bayes' rule to calculate this as
\begin{equation}
    p(M_1 \given \alignment{i}{}) = \frac{p(\alignment{i}{} \given
    M_1)p(M_1)}{p(\alignment{i}{} \given M_1)p(M_1) + p(\alignment{i}{} \given
    M_2)p(M_2)}.
    \label{eq:bayesmodelchoice}
\end{equation}
By comparing Equations \ref{eq:bayesrule} and \ref{eq:bayesmodelchoice}, we
see fundamental differences between Bayesian parameter estimation and
model choice.
In Equation \ref{eq:bayesrule}, we see that the posterior density of any
state $\myTheta{}$ of the model, is the prior density updated by the
probability of the data given the state $\myTheta{}$ (the likelihood of
$\myTheta{}$).
The integral over the entire parameter space of the model, which is defined by
the priors, only appears as a normalizing constant in the denominator.
Thus, as long as the prior distribution contains the values of $\myTheta{}$
that maximize the probability of the data (i.e., the maximum likelihood) and
the data are strongly informative relative to the prior, the values of the
parameters that maximize the posterior distribution will be relatively robust
to prior choice, even if the posterior distribution is technically incorrect
due to using the data to inform the priors.
This is why empirical Bayes can work well for estimating the values of model
parameters.

@@

However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
Bayesian model choice it is now the \emph{marginal} likelihood of a model that
updates the prior to yield the model's posterior probability.
Thus the integral over the entire parameter space of the likelihood weighted by
the prior probability density is no longer a normalizing constant, but rather
is what informs the posterior probability of that model from the data.
As a result, Bayesian model choice tends to be much more sensitive to priors
than parameter estimation.

Another important difference of Bayesian model choice illustrated by
Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the
posterior probability of a model, is not a function of \myTheta{}, because
it is integrated out of the marginal likelihoods of the candidate models.
Thus, unlike parameter estimates, the estimated posterior probability of a
model is a single value (rather than a distribution) lacking a measure of
posterior uncertainty.
Also, unlike model parameters, the posterior probabilities of candidate
models have no clear true values.
In some simple situations we might suspect that one model is correct and the
rest are wrong, but in most applications, we know that all the models under
consideration are oversimplifications of reality.
Thus, model posterior probabilities are inherently measures of our belief in
the models after our prior beliefs are updated by the data being analyzed.

This complicates the meaning of model posterior probabilities when Bayes' rule
is violated by informing priors with the same data to be analyzed.
For empirical Bayesian parameter estimation, the objective is that by giving the
data more weight relative to the prior, the posterior distribution will be
peaked near the true value(s) of the model's parameter(s).
There is no such justification for empirical Bayesian model choice, because
there are no true values for the model probabilities being estimated.
By using the data twice, we fail to account for prior uncertainty and mislead
our posterior beliefs in the models being compared; we will be over confident
in some models and under confident in others.

These distinctions between Bayesian parameter estimation and model choice can
be illustrated with a simple example.
Let us say we are interested in the fairness of a particular coin, and we
denote the unknown probability of it landing heads as \myTheta{}.
More specifically, we are interested in the probability of two models, $M_1$
and $M_2$.
In both models the outcomes of flipping the coin are assumed to be binomially
distributed, but under $M_1$ the coin is weighted toward landing heads (i.e.,
$\myTheta{} > 0.5)$), whereas under $M_2$, the coin is weighted toward landing
tails (i.e., $\myTheta{} < 0.5$).
We already have data from flipping a different coin 20 times that landed both
heads and tails 10 times, and so we decide to use these data in specifying a
beta prior on fairness of the new coin of $beta(a=10, b=10)$
(Figure~\ref{figCoinFlip}).
We collect data by flipping the coin $N=10$ times, $y=3$ of which land heads.
Given the beta distribution is a conjugate prior for a binomial likelihood, the
posterior distribution has the nice analytical form $\theta \given y,N \sim
beta(a + y, b + N - y)$, which for the new dataset is simply $beta(13, 17)$
(Figure~\ref{figCoinFlip}).
The maximum a posteriori (MAP) estimate of the probability of heads is 0.429,
and following Equation~\ref{eq:marginallikelhiood} the marginal likelihoods of
our models of interest are
\begin{equation}
    p(y=3, N=10 \given M_1) = \int_{0.5}^{1} p(y=3, N=10 \given
    \myTheta{}, M_1)p(\myTheta{} \given M_1) d\myTheta{} \approx 0.029,
\end{equation}
and
\begin{equation}
    p(y=3, N=10 \given M_2) = \int_{0}^{0.5} p(y=3, N=10 \given
    \myTheta{}, M_2)p(\myTheta{} \given M_2) d\myTheta{} \approx 0.097.
\end{equation}
Given the models have equal probability under our prior, and following
Equation~\ref{eq:bayesmodelchoice}, we can calculate the posterior
probability of Model 1 as
\begin{equation}
    p(M_1 \given y=3, N=10) = \frac{p(y=3, N=10 \given M_1)}{p(y=3, N=10 \given
    M_1) + p(y=3, N=10 \given M_2)} \approx 0.23.
\end{equation}
This is the correct posterior probability of Model 1 given our prior information
and data.

To give the data more weight relative to the prior, we could use it twice, and
calculate an empirical Bayes estimate using a prior of $beta(13,17)$.
This results in a distribution of $beta(16, 24)$ (Figure~\ref{figCoinFlip}),
with a MAP estimate of 0.395, and $p(M_1 \given y=3, N=10) = 0.10$.
The estimated posterior distribution of the parameter, and resulting MAP
estimate, is similar whether or not an empirically informed prior is used.
However, the posterior probability of Model 1 is very sensitive to the
empirical prior, decreasing by 56\%.
By using the empirically informed prior, we ignored prior uncertainty, leading
to an underestimate of our posterior uncertainty (Figure~\ref{figCoinFlip}).
While this did not greatly affect our estimate of \myTheta{}, it did mislead us
toward overconfidence in Model 2.
% Our mode estimate of \myTheta{} under the empirical prior might be closer
% to the true value than our posterior mode.
% Let us assume the truth of the matter is that the coin under study has a broad
% flat edge, and as a result lands on its edge at a certain frequency.
% Hence, both of our models are incorrect (i.e., the outcomes of flipping the
% coin are not binomially distributed).

@@

% \cwlnote{}{It seems like this section can be boiled down to ``empirical Bayes
%     approaches for model choice when using uniform priors can exclude the
%     $truth$ from the prior, which is a violation of Bayes theorem" or something
%     like that. Picking apart what they calculated and we calculated seems like
%     too much detail. I would try to simplify how this is being presented. This
%     can lead into how the use of more appropriate prior distributions should be
%     used.}
In addition to the theoretical concerns discussed above, there are practical
problems with using narrow, empirically informed, uniform priors for a method
that has already been shown to be biased toward models with less parameter
space.
The results of Hickerson et al.'s \citeyear{Hickerson2013} reanalysis of the
Philippine dataset strongly favored models with the narrowest, empirically
informed prior on divergence time, and thus their model-averaged posterior
estimates are dominated by models $M_1$ and $M_2$ (see Table 1 of
\citet{Hickerson2013}).
This is concerning, because the narrowest \divt{} prior used by
\citet{Hickerson2013} ($\divt{} \sim U(0,0.1)$) likely excludes the true
divergence times for at least some of the Philippine taxa, a major problem when
using uniform priors.
\citet{Hickerson2013} set this prior to match the 95\% highest posterior
density (HPD) interval for the mean divergence time estimated under one of the
priors used by \citet{Oaks2012} (see Tables 2 and 3 of \citet{Oaks2012}).
Given this interval estimate is for the \emph{mean} divergence time across all
22 taxa, it may be inappropriate to set this as the limit on the prior, because
many of the taxon pairs are expected to have diverged at times older than the
upper limit.
Furthermore, this prior is \emph{excluded} from the 95\% HPD interval estimates
of the mean divergence time under the other two priors explored by
\citet{Oaks2012} (under these priors the 95\% HPD is approximately 0.3--0.6;
see Table~6 of \citet{Oaks2012}).
Thus, the results of \citet{Hickerson2013} indicate that integration over
parameter space with low likelihood is biasing the method toward models with
smaller parameter space, and as a consequence is biasing the method toward
models that exclude the truth (i.e., toward estimating model-averaged
posteriors dominated by models that exclude true values of parameters of the
model).
